---
title: "Interest Rate vs Ratio - CFF3"
author: "UC Team"
format: html
editor: visual
---

```{r}
#| include: false

knitr::opts_chunk$set(
  fig.align = "center"
)

library(tidyverse)
library(lmtest)
library(tseries)
library(data.table)
library(dplyr)
library(tidyverse)
library(rmarkdown)
library(gapminder)
library(janitor)
library(lubridate)
library(scales)
library(gt)
library(patchwork)
library(kableExtra)
library(data.table)
library(corrplot)
library(zoo)
library(forecast)
library(fable)
library(feasts)
library(tsibble)
library(tseries)
library(ggplot2)
library(psych)
library(zoo)
library(gridExtra)
library(lmtest)
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

CFF_3 <- fread("C:\\Users\\varsh\\OneDrive\\Documents\\Graduate Case Studies\\Data\\CFF_3.csv", header = T)
#model_dates<-read.csv("Data/Model Dates.csv")
# Aggregating CFF_1 on month level (Difference of Actuals & Predicted)
CFF_3_agg <- CFF_3 %>% group_by(`Model Date ID`, `Month Number`, `Transaction Type`) %>% summarize(Amount = sum(Amount))
CFF_3_agg_wider <- CFF_3_agg %>%
  pivot_wider(names_from = `Transaction Type`,values_from = Amount) %>%
  mutate(Totalratio= A/P)

CFF_3_agg_ratio <- CFF_3_agg_wider[,c(2,5)]

CFF_3_agg_ratio<-as.data.frame(CFF_3_agg_ratio)
base_date <- as.Date("2013-10-01")
Date <- base_date+ months(CFF_3_agg_ratio$`Month Number`-1)
CFF_3_agg_ratio <- data.frame(Date = Date,ratio=CFF_3_agg_ratio$Totalratio)
CFF_3_agg_ratio

#Interest rate data
interest <- read.csv("C:\\Users\\varsh\\OneDrive\\Documents\\Graduate Case Studies\\Data\\FEDFUNDS_interest rates.csv")
interest <- interest %>%
  mutate(DATE = as.Date(as.character(DATE))) 
colnames(interest) <- c('Date','interest')

CFF_3_Interest_ratio <- merge(CFF_3_agg_ratio,interest, by = "Date")
CFF_3_Interest_ratio
write.csv(CFF_3_Interest_ratio,"CFF_3_Interest_ratio.csv")


```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

# Reading data
data <- fread("CFF_3_Interest_ratio.csv", header = T, fill = TRUE)

#head(data)
#summary(data)

data$V1 <- NULL
# Rename columns
colnames(data) <- c("time", "ratio", "irate")
#head(data)

datanew <- data %>%
   mutate(time = yearmonth(time))

head(datanew)
```

## Time Series plot of ratio of actual to predicted - Variable Y

The ratio plot has a lot of fluctuations in the time series plot with a steep dip in 2017.

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#Time series plot for ratio
datanew %>%
    ggplot() +
    geom_line(aes(time,ratio)) +
    theme_bw() +
    ggtitle("Time series plot of Ratio") +
    xlab("Time") +
    ylab("Ratio of actuals to predicted")
```

***The data is stationery .***

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
# Calculate summary statistics
#head(datanew)

# ADF Test for y1
invisible(adf.test(datanew$ratio))
#Non stationery because p value is greater than 0.05
# KPSS Test for y1
invisible(kpss.test(datanew$ratio))
#Non stationery because p values <<0.05
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
#Checking for mean stationarity for ratio column 
data_roll <- datanew %>%
  mutate(
    close_mean = zoo::rollmean(
      ratio, 
      k = 12, 
      fill = NA),
    close_sd = zoo::rollapply(
      ratio,  
      FUN = sd, 
      width = 12, 
      fill = NA)
  )

data_rollmean <- data_roll %>%
  ggplot() +
  geom_line(aes(time, ratio)) +
  geom_line(aes(time, close_mean),color='blue') +
  theme_bw() +
  ggtitle("Mean of Difference  over time") +
  ylab("Ratio") +
  xlab("Time")
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
data_roll <- datanew %>%
  mutate(
    close_mean = zoo::rollmean(
      ratio, 
      k = 12, 
      fill = NA),
    close_sd = zoo::rollapply(
      ratio, 
      FUN = sd, 
      width = 12, 
      fill = NA)
  )
data_rollsd <- data_roll %>%
  ggplot() +
  geom_line(aes(time, close_sd)) +
  geom_smooth(aes(time,close_sd),method='lm',se=F)+
  theme_bw() +
  ggtitle("Rolling SD Ratio of actuals and predicted") +
  ylab("Ratio of actuals and predicted") +
  xlab("Year")
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
data_trans <- datanew %>%
  mutate(data_log = log1p(ratio)) 
  
#data_trans %>%
  #ggplot() +
  #geom_line(aes(time, data_log)) +
  #geom_line(aes(time, data_boxcox),color='blue') +
  #theme_bw() +
  #ggtitle("Transformed ratio data") +
  #ylab("ratio data transformed") +
  #xlab("Year")
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
data_trans = datanew %>%
  mutate(data_log = log1p(ratio),
  value_log_diff = data_log - lag(data_log))

data_trans %>%
  ggplot() +
  geom_line(aes(time, value_log_diff),color = 'blue') +
  theme_bw() +
  ggtitle("Transformed and differenced ratio data") +
  ylab("Ratio data transformed and differenced") +
  xlab("Year")
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

invisible(adf.test(na.omit(data_trans$value_log_diff)))
```

## Time Series plot of Variable X : Interest Rate

Interest rates saw a steady incline from 2016, and has increased post that.

```{r}
#| echo: true
#| code-fold: true
#| warning: false


#Time series plot for Interest
datanew %>%
    ggplot() +
    geom_line(aes(time,irate)) +
    theme_bw() +
    ggtitle("Time series plot of interest rate") +
    xlab("Time") +
    ylab("Interest rate")
```

***Variable X is "non-stationary".***

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

# ADF Test for y2
adf.test(datanew$irate)
#Non stationery because p value is wayy greater than 0.05

# KPSS Test for y2
kpss.test(datanew$irate)
#Non stationery because p values <<0.05
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
#Checking for mean stationarity for Mortgage column 
data_roll <- datanew %>%
  mutate(
    close_mean_mrate = zoo::rollmean(
      irate, 
      k = 12, 
      fill = NA),
    close_sd_mrate = zoo::rollapply(
      irate, 
      FUN = sd, 
      width = 12, 
      fill = NA)
  )

data_rollmean_mrate <- data_roll %>%
  ggplot() +
  geom_line(aes(time, irate)) +
  geom_line(aes(time, close_mean_mrate),color='blue') +
  theme_bw() +
  ggtitle("Mean of interest Rate over time") +
  ylab("Interest Rate") +
  xlab("Time")
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
data_roll <- datanew %>%
  mutate(
    close_mean_mrate = zoo::rollmean(
      irate, 
      k = 12, 
      fill = NA),
    close_sd_mrate = zoo::rollapply(
      irate, 
      FUN = sd, 
      width = 12, 
      fill = NA)
  )
data_rollsd <- data_roll %>%
  ggplot() +
  geom_line(aes(time, close_sd_mrate)) +
  geom_smooth(aes(time,close_sd_mrate),method = 'lm',se = F) +
  theme_bw() +
  ggtitle("Rolling SD of Interest Rate") +
  ylab("Interest Rate") +
  xlab("Year")

#data_rollsd
```

#### *wt=log(xt)*

```{r}
#| echo: true
#| code-fold: true
#| warning: false
data_trans_2 <- datanew %>%
  mutate(data_log_irate = log1p(irate)) 
  #mutate(data_boxcox_unemp = forecast::BoxCox(ratio, lambda = "auto"))
  
#data_trans_2 %>%
 # ggplot() +
  #geom_line(aes(time, data_log_irate)) +
  #geom_line(aes(time, data_boxcox_unemp),color='blue') +
  #theme_bw() +
  #ggtitle("Transformed interest rate data") +
  #ylab("Interest rate data transformed") +
  #xlab("Year")
```

#### *wd=log(xt)-log(xt-1)*

```{r}
#| echo: true
#| code-fold: true
#| warning: false


data_trans_2 = data_trans_2 %>%
  mutate(data_log_irate = log1p(irate),
  value_log_diff_irate = data_log_irate - lag(data_log_irate))

data_trans_2 %>%
  ggplot() +
  geom_line(aes(time, value_log_diff_irate),color = 'blue') +
  theme_bw() +
  ggtitle("Transformed and differenced interest Rate") +
  ylab("Interest rate data transformed and differenced") +
  xlab("Year")
```

***Variable x is now "stationary"***.

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

adf.test(na.omit(data_trans_2$value_log_diff_irate))
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false

data_final <- merge(data_trans, data_trans_2, by.x = 'time', by.y = 'time')
#head(data_final)
data_final$ratio.y <- NULL
data_final$irate.y <- NULL
#data_final$data_log.y <- NULL
#data_final$value_log_diff.y <- NULL
#head(data_final)
colnames(data_final) <- c("time", "ratio", "irate", "value_log","value_log_diff", "value_log_irate", "value_log_diff_irate")
data_final[1, "ratio"] <- NA
#head(data_final)
```

## Granger Causality, Model Summary and Inferences :

Interest rate: xt <br>Ratio of actual to predicted ratio: yt <br>*Since both these series are nonstationary we have transformed them first before the Granger Causality test.* <br> Δxt = log(xt) − log (xt-1) <br> Δyt = log(𝑦t) − log (𝑦t-1) <br> xh = Δxt <br> yh = Δyt <br>

```{r}
#| echo: true
#| code-fold: true
#| warning: false

#Granger causality test
#grangertest(value_log_diff ~ value_log_diff_irate, order = 4, data = data_final)

#Dynamic regression
yh <- embed(data_final$ratio,6)
xh <- embed(data_final$value_log_diff_irate,6)
mh <- lm(yh[,1] ~ yh[,2:6] + xh[,5])
summary(mh)

#ANOVA
mh2 <- lm(yh[,1] ~ yh[,2:6])
anova(mh, mh2)
```

***Therefore, we can conclude that*** interest rate is statistically significant.

In this case , **xh5** is statistically significant :

For every unit increase in **xh5**, the response variable that is **Ratio is estimated to increase by** 0.88169 **units**, holding other predictors constant.
