---
title: "Mortgage Rate - Revenue - CFF1"
author: "UC Team"
format: html
editor: visual
---

```{r}
#| include: false

knitr::opts_chunk$set(
  fig.align = "center"
)

library(tidyverse)
library(lmtest)
library(tseries)
library(data.table)
library(dplyr)
library(tidyverse)
library(rmarkdown)
library(gapminder)
library(janitor)
library(lubridate)
library(scales)
library(gt)
library(patchwork)
library(kableExtra)
library(data.table)
library(corrplot)
library(zoo)
library(forecast)
library(fable)
library(feasts)
library(tsibble)
library(tseries)
library(ggplot2)
library(psych)
library(zoo)
library(gridExtra)
library(lmtest)
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false


CFF_1 <- fread("C:\\Users\\varsh\\OneDrive\\Documents\\Graduate Case Studies\\Data\\CFF_new.csv", header = T)
# Aggregating CFF_1 on month level (Difference of Actuals & Predicted)
CFF_1_agg <- CFF_1 %>% group_by(`Model Date ID`, `Month Number`, `Transaction Type`) %>% summarize(Amount = sum(Amount))
CFF_1_agg_wider <- CFF_1_agg %>%
  pivot_wider(names_from = `Transaction Type`,values_from = Amount) %>%
  mutate(TotalRevenue= A)

CFF_1_agg_revenue <- CFF_1_agg_wider[,c(2,5)]

CFF_1_agg_revenue<-as.data.frame(CFF_1_agg_revenue)
base_date <- as.Date("2007-10-01")
Date <- base_date+ months(CFF_1_agg_revenue$`Month Number`-1)
CFF_1_agg_revenue <- data.frame(Date = Date,Revenue=CFF_1_agg_revenue$TotalRevenue)
#CFF_1_agg_revenue
# CFF_1_agg_revenue Final file for CFF_1 monthly aggregated data national level 


mortgage_rate<-read.csv("C:\\Users\\varsh\\OneDrive\\Documents\\Graduate Case Studies\\Data\\WeeklyMortgageRate_edited.csv ")
#head(mortgage_rate)
#colnames(mortgage_rate) <- c('Date','Mrate')





names(mortgage_rate)[2] <- "Mrate"

mortgage_rate <- mortgage_rate %>%
  filter(Year>=2007 & Date_Key >= '2007-10') %>% # Corrected Date_Key format
  mutate(Week = yearmonth(Date_Key)) %>% 
  group_by(Week) %>%
  summarise(Mrate = mean(Mrate))
names(mortgage_rate)[1] <- "Date"
#head(mortgage_rate)

#Appending unemployment rate to CFF_1 Ratio file
CFF_1_Revenue_Mortgage<-merge(CFF_1_agg_revenue,mortgage_rate, by ="Date")
head(CFF_1_Revenue_Mortgage)
#write.csv(CFF_1_Revenue_Mortgage,"CFF_1_Revenue_Mortgage.csv")
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
# Reading data
data <- fread("CFF_1_Revenue_Mortgage.csv", header = T, fill = TRUE)
#head(data)
#summary(data)

data$V1 <- NULL
# Rename columns
colnames(data) <- c("time", "Revenue", "Mrate")
#head(data)

datanew <- data %>%
   mutate(time = yearmonth(time))

#head(datanew)
```

## Time Series plot of Revenue of actuals and projections - Variable Y

The Revenue plot has a lot of fluctuations in the time series plot with a peak right before the start of the year 2012 and mostly follows an exponentially decreasing trend.

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#Time series plot for Revenue
datanew %>%
    ggplot() +
    geom_line(aes(time,Revenue)) +
    theme_bw() +
    ggtitle("Time series plot of Revenue") +
    xlab("time") +
    ylab("Revenue")

```

***The data is not stationery - does not pass the stationairty tests .***

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
# Calculate summary statistics
#head(datanew)

# ADF Test for y1
invisible(adf.test(datanew$Revenue))
#Non stationery because p value is greater than 0.05
# KPSS Test for y1
invisible(kpss.test(datanew$Revenue))
#Non stationery because p values <<0.05
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
#Checking for mean stationarity for Revenue column 
data_roll <- datanew %>%
  mutate(
    close_mean = zoo::rollmean(
      Revenue, 
      k = 12, 
      fill = NA),
    close_sd = zoo::rollapply(
      Revenue,  
      FUN = sd, 
      width = 12, 
      fill = NA)
  )

data_rollmean <- data_roll %>%
  ggplot() +
  geom_line(aes(time, Revenue)) +
  geom_line(aes(time, close_mean),color='blue') +
  theme_bw() +
  ggtitle("Mean of Difference  over time") +
  ylab("Revenue of  actuals ") +
  xlab("Time")

```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
data_roll <- datanew %>%
  mutate(
    close_mean = zoo::rollmean(
      Revenue, 
      k = 12, 
      fill = NA),
    close_sd = zoo::rollapply(
      Revenue, 
      FUN = sd, 
      width = 12, 
      fill = NA)
  )
data_rollsd <- data_roll %>%
  ggplot() +
  geom_line(aes(time, close_sd)) +
  geom_smooth(aes(time,close_sd),method='lm',se=F)+
  theme_bw() +
  ggtitle("Rolling SD Revenue of actuals and predicted") +
  ylab("Revenue of actuals and predicted") +
  xlab("Year")



```

#### [*wt=log(yt)*]{.underline}

```{r}
#| echo: true
#| code-fold: true
#| warning: false
data_trans <- datanew %>%
  mutate(data_log = log1p(Revenue)) 
  

#data_trans %>%
  #ggplot() +
  #geom_line(aes(time, data_log)) +
  #geom_line(aes(time, data_boxcox),color='blue') +
  #theme_bw() +
  #ggtitle("Transformed Revenue data") +
  #ylab("Revenue data transformed") +
  #xlab("Year")
```

**Treating for Mean Non Stationarity for Variable Y**

Considering the first difference for to make it mean stationery as follows : ***wd=log(y­~t~)-log(y~t-1~)***

```{r}
#| echo: true
#| code-fold: true
#| warning: false
data_trans = datanew %>%
  mutate(data_log = log1p(Revenue),
  value_log_diff = data_log - lag(data_log))

data_trans %>%
  ggplot() +
  geom_line(aes(time, value_log_diff),color = 'blue') +
  theme_bw() +
  ggtitle("Transformed and differenced Revenue data") +
  ylab("Revenue data transformed and differenced") +
  xlab("Year")
```

### Stationarity Check for Variable y :

***Variable y is now "stationary"***.

*p value is less than 0.05*

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

invisible(adf.test(na.omit(data_trans$value_log_diff)))
```

## Time Series plot of Variable X : Mortgage rate

Mortgage rate follows an decreasing curve after the initial high in 2008.

```{r}
#| echo: true
#| code-fold: true
#| warning: false


#Time series plot for Mortgage rate
datanew %>%
    ggplot() +
    geom_line(aes(time,Mrate)) +
    theme_bw() +
    ggtitle("Time series plot of Mortgage rate") +
    xlab("time") +
    ylab(" Mortgage rate")
```

### Stationarity Test for Variable X :

***Variable X is "non-stationary".***

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

# ADF Test for y2
adf.test(datanew$Mrate)
#Non stationery because p value is wayy greater than 0.05

# KPSS Test for y2
kpss.test(datanew$Mrate)
#Non stationery because p values <<0.05
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#Checking for mean stationarity for Mortgage column 
data_roll <- datanew %>%
  mutate(
    close_mean_mrate = zoo::rollmean(
      Mrate, 
      k = 12, 
      fill = NA),
    close_sd_mrate = zoo::rollapply(
      Mrate, 
      FUN = sd, 
      width = 12, 
      fill = NA)
  )

data_rollmean_mrate <- data_roll %>%
  ggplot() +
  geom_line(aes(time, Mrate)) +
  geom_line(aes(time, close_mean_mrate),color='blue') +
  theme_bw() +
  ggtitle("Mean of Mortgage Rate  over time") +
  ylab("Mortgage Rate") +
  xlab("Time")

```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
data_roll <- datanew %>%
  mutate(
    close_mean_mrate = zoo::rollmean(
      Mrate, 
      k = 12, 
      fill = NA),
    close_sd_mrate = zoo::rollapply(
      Mrate, 
      FUN = sd, 
      width = 12, 
      fill = NA)
  )
data_rollsd <- data_roll %>%
  ggplot() +
  geom_line(aes(time, close_sd_mrate)) +
  geom_smooth(aes(time,close_sd_mrate),method='lm',se=F)+
  theme_bw() +
  ggtitle("Rolling SD of Mortgage Rate") +
  ylab("Mortgage Rate") +
  xlab("Year")

#data_rollsd
```

#### [*wt=log(xt)*]{.underline}

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
data_trans_2 <- datanew %>%
  mutate(data_log_mrate = log1p(Mrate)) 
  #mutate(data_boxcox_unemp = forecast::BoxCox(ratio, lambda = "auto"))
  

data_trans_2 %>%
  ggplot() +
  geom_line(aes(time, data_log_mrate)) +
  #geom_line(aes(time, data_boxcox_unemp),color='blue') +
  theme_bw() +
  ggtitle("Transformed mortgage rate data") +
  ylab("Mortgage data transformed") +
  xlab("Year")
```

***wd=log(xt)-log(xt-1)***

```{r}
#| echo: true
#| code-fold: true
#| warning: false

data_trans_2 = data_trans_2 %>%
  mutate(data_log_mrate = log1p(Mrate),
  value_log_diff_mrate = data_log_mrate - lag(data_log_mrate))

data_trans_2 %>%
  ggplot() +
  geom_line(aes(time, value_log_diff_mrate),color = 'blue') +
  theme_bw() +
  ggtitle("Transformed and differenced mortgage Rate") +
  ylab("Mortgage data transformed and differenced") +
  xlab("Year")
```

***Variable x is now "stationary"***.

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false

adf.test(na.omit(data_trans_2$value_log_diff_mrate))
```

```{r}
#| echo: true
#| code-fold: true
#| warning: false
#| include: false
data_final <- merge(data_trans, data_trans_2, by.x = 'time', by.y = 'time')
#head(data_final)
data_final$Revenue.y <- NULL
data_final$Mrate.y <- NULL
#data_final$data_log.y <- NULL
#data_final$value_log_diff.y <- NULL
#head(data_final)
colnames(data_final) <- c("time", "revenue", "Mrate", "value_log","value_log_diff", "value_log_mrate", "value_log_diff_mrate")
#head(data_final)
```

### Granger Causality, Model Summary and Inferences :

```{r}
#| echo: true
#| code-fold: true
#| warning: false


#Granger causality test
#grangertest(value_log_diff ~ value_log_diff_mrate, order = 3, data = data_final)

#Dynamic regression
yh <- embed(data_final$value_log_diff,4)
xh <- embed(data_final$value_log_diff_mrate,4)
mh <- lm(yh[,1] ~ yh[,2:4] + xh)
summary(mh)

#ANOVA
mh2 <- lm(yh[,1] ~ yh[,2:4])
anova(mh, mh2)
```

***Therefore, we can conclude that*** because of the high p-value of the Anova test between model 1 and 2, we can conclude that Mortgage does not Granger-cause Revenue.
